### 2013-08-16 paper/paper.tex
Wrote the paper that describes mathematically the particle in a box

\documentclass[a4paper, 10pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\hypersetup{colorlinks=false}

\author{George Zakhour}
\title{{\bf Simulating a two dimensional particle in a square quantum box with
        CUDA}}
\date{\today}

\begin{document}

\maketitle
\newpage
\tableofcontents
\newpage

\section{Background, Motive and Experiment}
The particle in box problem is one of the first problems given to undergraduate
students in a course about quantum mechanics. It showcases how the energy
of particles is quantized and it highlights the probabilistic nature of
quantum mechanics, especially with the idea that the particle is not located in
a fixed position but it has a likelihood of existing at any point in space.\\\\
The particle in a box is an experiment in which a particle is stuck inside a
box and cannot escape. The energy outside this box is $\infty$ while inside it
is $0$. The particle thus moves inside the box of dimensions $L$x$L$.\\\\
Trying to imagine how these probabilities are scattered in the box is hard and one
can't do without a simulation of the physical properties of the particle
(position, energy and momentum).\\\\
In this simulation I try and simulate the "particle in a box" problem to
display the probabilities of the position and the energy of the particle at
each state.

\section{Expected Output}
Searching online, I have found a Youtube video that simulates the particle in a
box\footnote{ Particle in a Box - Youtube {\emph
http://www.youtube.com/watch?v=jevKmFfcaxE}}, but lacks essential information
on the state of the system, for example the dimensions of the box and the
energy levels of the particle at each frame. Although lacking these
essential variables I will base my results on the video.\\

\begin{figure}[hb]
    \centering
        \includegraphics[width=5cm]{graphics/online_simulation.png}
    \caption{A screenshot of a "particle in a box" simulation found on Youtube}
\end{figure}

\newpage

\section{The mathematics of a particle in a box}
    \subsection{In one dimension}
        \subsubsection{Finding the Wave Function from Schrodinger's Equation}
        \label{sec:wave_equation_oned}
        The time dependent Schrodinger equation is given as:
        \begin{equation} \label{eq:time_dependant_equation}
            \left ( \frac{-\hbar^2}{2m} \nabla^2 + V \right ) \Psi(x,t)
            = i\hbar\frac{\partial}{\partial t}\Psi(x, t)   
        \end{equation}
        The time independent Schrodinger equation is given as:
        \begin{equation} \label{eq:time_independent_equation}
            \left ( \frac{-\hbar^2}{2m} \nabla^2 + V \right ) \Phi(x)
            = E \Phi(x)
        \end{equation}
        Where $\Psi$ is the wave equation. In terms of $\Phi$, $\Psi$ is
        denoted as:
        \begin{equation} \label{eq:dep_in_equation}
            \Psi(x, t) = e^{-i(E/\hbar)t}\Phi(x)
        \end{equation}
        Assuming the following is a solution to (\ref{eq:time_independent_equation}):
        \begin{equation} \label{eq:assumed_solution}
            \Phi(x) = A\cos(kx) + B\sin(kx)
        \end{equation}
        And given these two conditions that arise from the experiment:
        $$ \Phi(0) = \Phi(L) = 0 $$
        We plug them in the (\ref{eq:assumed_solution}) and find the following:
        \begin{align}
            \left\{\begin{matrix}
            A & = & 0 \\ 
            k & = & \frac{n}{L}\pi & \text{(n is the energy level)}
            \end{matrix}\right.
        \end{align}
        To find the value of $B$ we need to normalize the equation.\\
        The probability of finding the particle inside $[0; L]$ is $1$ because
        it cannot escape.
        \begin{align}
            \int_{0}^{L} |\Phi|^2 dx &= 1  \\
            B^2\int_{0}^{L} \sin^2\left ( \frac{n}{L}\pi x \right ) dx &= 1
        \end{align}
        And thus $B = \sqrt{\frac{2}{L}}$.\\
        Finally we get the solution to the time independent Schrodinger equation
        \begin{align}
            \Phi(x) = \sqrt{\frac{2}{L}}\sin(\frac{n}{L}\pi x)
        \end{align}

        \subsubsection{Energy at each quantum level $n$}
        \label{sec:energy_oned}
        We note the following
        \begin{align}
            \frac{\partial^2 }{\partial x^2} \Phi = -\left (
            \frac{n\pi}{L}\right )^2 \Phi
        \end{align}
        If we replace the results in (\ref{eq:time_independent_equation}) we find:
        $$ E_n = \frac{\hbar^2 n^2 \pi^2}{ 2 m L^2} $$
        Or simply:
        \begin{align*}
            E_n &= n^2 E_1 \\
            E_1 &= \frac{\hbar^2 \pi^2}{2 m L^2}
        \end{align*}

        \subsubsection{Probability function of the position}
        \label{sec:position_oned}
        The probability of finding the particle between $a$ and $b$ is the following:
        \begin{equation} \label{eq:probability_position}
            \left\{\begin{matrix}
            \int_{a}^{b} |\Phi|^2 dx & \text{if} \; \: a,b \in [0, L] \\ 
            0 & \text{otherwise}
            \end{matrix}\right.
        \end{equation}
        If we wish to find the position inside a box of width $\epsilon$ and
        center $a$ we would integrate between $a-\epsilon/2$ and $a+\epsilon/2$
        where both ends are between $0$ and $L$\\
        The integration will lead to the following:
        $$ P(x) = \frac{\epsilon}{L} + \frac{1}{2n\pi} \left [ 
            \sin\left ( \frac{2n\pi}{L}(a - \epsilon/2)\right ) -
            \sin\left ( \frac{2n\pi}{L}(a + \epsilon/2)\right )
        \right ]$$
        Which can be reduced more to the following form:
        $$ P(x) = \frac{\epsilon}{L} - \frac{1}{n\pi}\sin\left( \frac{n\pi}{L}\epsilon \right)
        \cos\left(\frac{2n\pi}{L}x\right)$$

    \subsection{In two dimensions}
        \subsubsection{The time-independent solution}
        This time we suppose the solution is
        \begin{equation} \label{eq:suppose_twod_solution}
            \Phi(x, y) = X(x)Y(y)
        \end{equation}
        \begin{align*}
            X(x) &= A\cos(k_x x) + B \sin(k_x x) \\
            Y(y) &= C\cos(k_y y) + D \sin(k_y y) \\
        \end{align*}
        Plugging (\ref{eq:suppose_twod_solution}) in (\ref{eq:time_independent_equation})
        and doing similar operations as \emph{~\ref{sec:wave_equation_oned}} we get
        the following solution:
        \begin{equation} \label{eq:twod_solution}
            \Phi(x, y) = \frac{2}{L} \sin\left( \frac{n_x\pi}{L}x\right)
                         \sin\left( \frac{n_y\pi}{L} y\right)
        \end{equation}

        \subsubsection{Energy in two dimensions}
        Doing the same steps as \emph{~\ref{sec:energy_oned}} we find that the energy has
        become:
        \begin{align}
            E_{n_x, n_y} &= \frac{\hbar^2 \pi^2}{2mL^2} (n^2_x + n^2_y)
        \end{align}
        Or simply:
        \begin{align*}
        E_{n_x, n_y} &= (n^2_x + n^2_y) \; E_{1,1} \\
        E_{1,1} &= \frac{\hbar^2 \pi^2}{2mL^2} 
        \end{align*}

        \subsubsection{Probability function of the position}
        Similar to section \emph{~\ref{sec:position_oned}} to find the
        probability of finding the particle inside the box of size
        $\epsilon$x$\epsilon$ we integrate the wave function inside a
        box of center $(x,y)$ and width $\epsilon$.
        $$ P(x, y) = \int^{y+\epsilon/2}_{y-\epsilon/2}\int^{x+\epsilon/2}_{x-\epsilon/2}
          |\Phi(x, y)|^2 dx dy $$
        After evaluating the integral we get the following function:
        \begin{equation} \label{eq:position_twod_equation}
            P(x, y) = \frac{1}{L^2} \; p(x) \; p(y)
        \end{equation}
        $$
        p(\alpha) = \epsilon - \frac{L}{n_\alpha\pi} \cos\left( \frac{2n_\alpha\pi}{L}\alpha \right)
        \sin\left( \frac{n_\alpha\pi}{L} \epsilon\right)
        $$

        \subsection{Generalizing the time-independent solution}
        The equation that we have found so far depends on two quantum numbers
        $n_x$ and $n_y$ and describes the particle for these energy levels only.
        However the final time-dependant equation is a combination of many of these
        equations. The final time-independent equation is:
        \begin{equation} \label{eq:general_position}
            \Phi(x,y) = \sum_{n} c_n \Phi_{n_x, n_y}(x,y)
        \end{equation}
        Where the constant $c_n$ is the square root of the probability of getting
        the equation $\Phi_{n_x, n_y}$ that describes the particle in the box.

        \subsection{Generalizing the probability function}
        Since we have a more general wave function we need to have a more general
        probability function for the position. The new definition is generated from
        integrating $\Phi$ between $a-\epsilon/2$ and $a+\epsilon/2$. The result is:
        $$ P(x, y) = \sum_{n} c^2_n \: P_n(x, y)$$

\newpage

\section{Implementation}
    \subsection{Memory Management}
    \subsection{Optimizations}
    \subsection{Implementations Details}

\newpage
\section{Results}

\newpage
\section{Code}
    \subsection{Code}
\begin{verbatim}
\end{verbatim}
    \subsection{License: GNU General Public License}
    Copyright (C) 2013 George Zakhour\\

    This program is free software: you can redistribute it and/or modify it under
    the terms of the GNU General Public License as published by the Free Software
    Foundation, either version 3 of the License, or (at your option) any later
    version.\\

    This program is distributed in the hope that it will be useful, but WITHOUT ANY
    WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A
    PARTICULAR PURPOSE. See the GNU General Public License for more details.\\

    You should have received a copy of the GNU General Public License along with
    this program. If not, see http://www.gnu.org/licenses/.\\

\end{document}

### 2013-08-17 paper/paper.tex
Created probability and particle-creation functions in code Added documentation to these functions in the paper Added conventions used to the paper

@@
 \section{Implementation}
-    \subsection{Memory Management}
-    \subsection{Optimizations}
-    \subsection{Implementations Details}
+    \subsection{Conventions}
+    Some of the conventions adopted through the code are
+    \begin{itemize}
+        \item words in function and struct names are seperated by an \_ (underscore)
+        \item words in variable names are glued together with the first letter
+              capitalized. Except in the case of the first character.
+        \item functions that start with \_ are helper functions
+              and should not be used directly. They may be considered like private
+              methods and properties in object-oriented code. No major documentation
+              is provided for these functions and the author only uses inline code
+              comments to document them
+        \item Tiny mathematical and physical constants, such as $\hbar$ are expressed
+              as float numbers without their orders. For example $\hbar = 1.054\cdot10^{-34}$
+              is defined the following way \texttt{\#define HBAR 1.054 // without e-34}
+    \end{itemize}
+
+\newpage
+\section{API}
+    \subsection{Structs}
+        \subsubsection{\texttt{\_wave\_function}}
+        This struct is used by the \texttt{particle} struct and describes a wave function
+        \begin{itemize}
+            \item \texttt{int n} the energy level of the particle described by the wave function
+            \item \texttt{float energy} the energy of the particle at that energy level
+        \end{itemize}
+        \subsubsection{\texttt{particle}}
+        This struct represents a particle and have the following elements
+        \begin{itemize}
+            \item \texttt{float mass} the mass of the particle
+            \item \texttt{\_wave\_function* wave} the wave function that describes the state of
+                  the particle
+        \end{itemize}
+
+    \subsection{Functions}
+        \subsubsection{\texttt{create\_particle}}
+        This function is used to create a new particle.
+        \begin{verbatim}
+void create_particle(particle* p, int n, float mass);
+        \end{verbatim}
+        \begin{itemize}
+            \item \texttt{particle* p} the particle to create
+            \item \texttt{int n} the energy level of the particle
+            \item \texttt{float mass} the mass of the particle
+        \end{itemize}
+        \subsubsection{\texttt{probability}}
+        This function finds the probability of finding the particle in the box of center $(x, y)$
+        and of dimensions $\epsilon$x$\epsilon$\\
+        The function returns a \texttt{float}; the probability of finding the particle at points (x, y)
+        \begin{verbatim}
+float probability(particle *p, float x, float y);
+        \end{verbatim}
+        \begin{itemize}
+            \item \texttt{particle* p} the particle to find the probability of
+            \item \texttt{float x} the x-coordinate of the center of the box 
+            \item \texttt{float y} the y-coordinate of the center of the box
+        \end{itemize}
 

### 2013-08-17 src/pbox.cu

@@ -1,0 +1,89 @@
+#include <stdio.h>
+#include <cuda.h>
+#include <math.h>
+
+
+#define PI   3.1415
+#define L    1.0          // the length of the box
+#define DIM  500          // the dimensions of the window
+#define E    (L/DIM)
+#define HBAR 1.054        // without the e-34
+
+#define MAX(a,b) a > b ? a : b
+#define MIN(a,b) a < b ? a : b
+
+
+typedef struct _wave_function {
+    int n;          // the energy level of the wave function
+    float energy;   // the energy of the particle at that level
+} _wave_function;
+
+typedef struct particle {
+    float mass;           // the mass of the particle
+    _wave_function *wave; // the wave function of the particle 
+} particle;
+
+
+/**
+  *Create a new particle
+ *
+  *@param particle      the particle to create
+  *@param int           the energy level of the particle
+  *@param float         the mass of the particle
+*/
+void create_particle(particle *p, int n, float mass) {
+    p->mass = mass;
+    p->wave = (_wave_function *) malloc(sizeof(_wave_function));
+    p->wave->n = n;
+    p->wave->energy = HBAR*HBAR * PI*PI * n*n / (mass * L*L);
+}
+
+/**
+  *Get the probability of finding a particle in one dimension
+ *
+  *@param particle      the particle to find the probability of
+  *@param float         the position in which to find the probability at
+  *@return              the probability of finding the particle at the point x
+*/
+float _probability_1d(particle *p, float x) {
+    int n   =  p->wave->n;
+    float s =  E/L - (1/(n*PI)) * cos(2*n*PI*x/L) * sin(n*PI / DIM);
+    return s;
+}
+
+/**
+  *Get the probability of finding a particle in two dimensions
+ *
+  *@param particle      the particle to find the probability of
+  *@param float         the x-position in which to find the probability at
+  *@param float         the y-position in which to find the probability at
+  *@return              the probability of finding the particle at the point (x, y)
+*/
+float probability(particle *p, float x, float y) {
+    if(x > 0 && x < L && y > 0 && y < L) {
+        return 1.0/(L*L) * _probability_1d(p, x) * _probability_1d(p, y);
+    }
+    return 0;
+}
+
+int main(int argc, const char *argv[]) { 
+
+    particle p;
+    create_particle(&p, 3, 2.5);
+
[…]

### 2013-08-17 src/pbox.cu
Added a new function _create_wave to seperate the code more

@@ -24,26 +24,25 @@ typedef struct particle {
 } particle;
 
 
+
 /**
-  *Create a new particle
+ * Create a new wave
  *
-  *@param particle      the particle to create
-  *@param int           the energy level of the particle
-  *@param float         the mass of the particle
+ * @param _wave_function  the wave function to create
+ * @param int             the energy level of the particle
+ * @param float           the mass of the particle
 */
-void create_particle(particle *p, int n, float mass) {
-    p->mass = mass;
-    p->wave = (_wave_function *) malloc(sizeof(_wave_function));
-    p->wave->n = n;
-    p->wave->energy = HBAR*HBAR * PI*PI * n*n / (mass * L*L);
+void _create_wave(_wave_function *wave, int n, float mass) {
+    wave->n = n;
+    wave->energy = HBAR*HBAR * PI*PI * n*n / (mass * L*L);
 }
 
 /**
-  *Get the probability of finding a particle in one dimension
+ * Get the probability of finding a particle in one dimension
  *
-  *@param particle      the particle to find the probability of
-  *@param float         the position in which to find the probability at
-  *@return              the probability of finding the particle at the point x
+ * @param particle      the particle to find the probability of
+ * @param float         the position in which to find the probability at
+ * @return              the probability of finding the particle at the point x
 */
 float _probability_1d(particle *p, float x) {
     int n   =  p->wave->n;
@@ -51,13 +50,28 @@ float _probability_1d(particle *p, float x) {
     return s;
 }
 
+
+
 /**
-  *Get the probability of finding a particle in two dimensions
+ * Create a new particle
+ * 
+ * @param particle      the particle to create
+ * @param int           the energy level of the particle
+ * @param float         the mass of the particle
+*/
+void create_particle(particle *p, int n, float mass) {
+    p->mass = mass;
+    p->wave = (_wave_function *) malloc(sizeof(_wave_function));
+    _create_wave(p->wave, n, mass);
+}
+
+/**
+ * Get the probability of finding a particle in two dimensions
  *
-  *@param particle      the particle to find the probability of
-  *@param float         the x-position in which to find the probability at
-  *@param float         the y-position in which to find the probability at
-  *@return              the probability of finding the particle at the point (x, y)
+ * @param particle      the particle to find the probability of
+ * @param float         the x-position in which to find the probability at
+ * @param float         the y-position in which to find the probability at
[…]

### 2013-08-18 src/pbox.cu
Added support for multiple wave equations with each's probability

@@ -6,7 +6,7 @@
 #define PI   3.1415
 #define L    1.0          // the length of the box
 #define DIM  500          // the dimensions of the window
-#define E    (L/DIM)
+#define E    (L/DIM)      // epsilon
 #define HBAR 1.054        // without the e-34
 
 #define MAX(a,b) a > b ? a : b
@@ -14,13 +14,15 @@
 
 
 typedef struct _wave_function {
-    int n;          // the energy level of the wave function
-    float energy;   // the energy of the particle at that level
+    int n;             // the energy level of the wave function
+    float energy;      // the energy of the particle at that level
+    float probability; // the probability of getting this wave function (c^2_n)
 } _wave_function;
 
 typedef struct particle {
-    float mass;           // the mass of the particle
-    _wave_function *wave; // the wave function of the particle 
+    float mass;            // the mass of the particle
+    _wave_function **wave; // array of wave functions for the particle 
+    int waveCount;
 } particle;
 
 
@@ -31,10 +33,12 @@ typedef struct particle {
  * @param _wave_function  the wave function to create
  * @param int             the energy level of the particle
  * @param float           the mass of the particle
+ * @param float           the probability of the wave function (c^2_n)
 */
-void _create_wave(_wave_function *wave, int n, float mass) {
+void _create_wave(_wave_function *wave, int n, float mass, float p) {
     wave->n = n;
     wave->energy = HBAR*HBAR * PI*PI * n*n / (mass * L*L);
+    wave->probability = p;
 }
 
 /**
@@ -44,8 +48,8 @@ void _create_wave(_wave_function *wave, int n, float mass) {
  * @param float         the position in which to find the probability at
  * @return              the probability of finding the particle at the point x
 */
-float _probability_1d(particle *p, float x) {
-    int n   =  p->wave->n;
+float _probability_1d(_wave_function *w, float x) {
+    int n   =  w->n;
     float s =  E/L - (1/(n*PI)) * cos(2*n*PI*x/L) * sin(n*PI / DIM);
     return s;
 }
@@ -56,13 +60,18 @@ float _probability_1d(particle *p, float x) {
  * Create a new particle
  * 
  * @param particle      the particle to create
- * @param int           the energy level of the particle
[…]

### 2013-08-18 src/pbox.cu
If probabilities != 0, the program raises and error and quits

@@ -11,6 +11,7 @@
 
 #define MAX(a,b) a > b ? a : b
 #define MIN(a,b) a < b ? a : b
+#define ERROR(msg) do { fprintf(stderr, "%s(%d): %s\n", __FILE__, __LINE__, msg); exit(0); } while(0);
 
 
 typedef struct _wave_function {
@@ -65,6 +66,10 @@ float _probability_1d(_wave_function *w, float x) {
  * @param float         the mass of the particle
 */
 void create_particle(particle *p, int N, float *probabilities, float mass) {
+    float sumProba = 0;
+    for(int i=0;i<N;i++) sumProba += probabilities[i];
+    if(sumProba != 1)
+        ERROR("Sum of all the probabilities is not 1.");
     p->mass = mass;
     p->waveCount = N;
     p->wave = (_wave_function **) malloc(N*sizeof(_wave_function));

### 2013-08-18 src/pbox.cu
Added a next_probability function that generates new c^2_n

@@ -56,6 +56,28 @@ float _probability_1d(_wave_function *w, float x) {
 }
 
 
+/**
+ * Given numbers, the function generates the next set of probabilities
+ * based on these numbers.
+ *
+ * @param int           the numbers to generate the probabilities from
+ * @param int           the length of the numbers array (should be the same
+                        for the probabilities array);
+ * @param float         array of probabilities to write to
+*/
+void next_probabilities(int *numbers, int N, float *probabilities) {
+    int sum = 0;
+    int carry = 1;
+    for(int i=N-1;i>=0;i--) {
+        numbers[i] +=  carry;
+        carry       =  numbers[i] / 10;
+        numbers[i] %=  10;
+        sum        +=  numbers[i];
+    }
+    for(int i=0;i<N;i++) {
+        probabilities[i] = 1.0f*numbers[i] / sum;
+    }
+}
 
 /**
  * Create a new particle
@@ -66,10 +88,6 @@ float _probability_1d(_wave_function *w, float x) {
  * @param float         the mass of the particle
 */
 void create_particle(particle *p, int N, float *probabilities, float mass) {
-    float sumProba = 0;
-    for(int i=0;i<N;i++) sumProba += probabilities[i];
-    if(sumProba != 1)
-        ERROR("Sum of all the probabilities is not 1.");
     p->mass = mass;
     p->waveCount = N;
     p->wave = (_wave_function **) malloc(N*sizeof(_wave_function));
@@ -106,10 +124,9 @@ int main(int argc, const char *argv[]) {
     int N = 10;
     float *pro = (float*) malloc(N*sizeof(float));
     for(int i=0;i<N;i++) pro[i] = 0.1f;
-    pro[6] = 0.05f;
-    pro[7] = 0.15f;
     create_particle(&p, N, pro, 2.5);
 
+    printf("Testing on a particle\n");
     printf("The particle has mass %f and has %d energy levels\n",
            p.mass, p.waveCount);
     for(int i=0;i<p.waveCount;i++)
@@ -137,5 +154,16 @@ int main(int argc, const char *argv[]) {
     printf("The probability of finding the particle in (%f, %f) is %.5f%% or %.3f%% of the maximum\n",
            x, y, probability(&p, x, y)*100, probability(&p, x, y)*100/max_proba);
 
+    printf("Testing on generating new probabilities\n");
[…]

### 2013-08-20 src/pbox.cu
Go back in case of huge error

@@ -5,13 +5,16 @@
 
 #define PI   3.1415
 #define L    1.0          // the length of the box
-#define DIM  500          // the dimensions of the window
+#define DIM  512          // the dimensions of the window (should be 2^n)
 #define E    (L/DIM)      // epsilon
 #define HBAR 1.054        // without the e-34
 
 #define MAX(a,b) a > b ? a : b
 #define MIN(a,b) a < b ? a : b
-#define ERROR(msg) do { fprintf(stderr, "%s(%d): %s\n", __FILE__, __LINE__, msg); exit(0); } while(0);
+#define ERROR(msg) do {                                               \
+            fprintf(stderr, "%s(%d): %s\n", __FILE__, __LINE__, msg); \
+            exit(0);                                                  \
+        } while(0);
 
 
 typedef struct _wave_function {
@@ -22,7 +25,6 @@ typedef struct _wave_function {
 
 typedef struct particle {
     float mass;            // the mass of the particle
-    _wave_function **wave; // array of wave functions for the particle 
     int waveCount;
 } particle;
 
@@ -42,30 +44,17 @@ void _create_wave(_wave_function *wave, int n, float mass, float p) {
     wave->probability = p;
 }
 
-/**
- * Get the probability of finding a particle in one dimension
- *
- * @param particle      the particle to find the probability of
- * @param float         the position in which to find the probability at
- * @return              the probability of finding the particle at the point x
-*/
-float _probability_1d(_wave_function *w, float x) {
-    int n   =  w->n;
-    float s =  E/L - (1/(n*PI)) * cos(2*n*PI*x/L) * sin(n*PI / DIM);
-    return s;
-}
-
 
 /**
  * Given numbers, the function generates the next set of probabilities
  * based on these numbers.
  *
- * @param int           the numbers to generate the probabilities from
- * @param int           the length of the numbers array (should be the same
-                        for the probabilities array);
- * @param float         array of probabilities to write to
+ * @param int            the numbers to generate the probabilities from
+ * @param int            the length of the numbers array (should be the same
+                         for the probabilities array);
[…]

### 2013-08-20 src/pbox.cu
A working prototype without a GUI

@@ -9,6 +9,7 @@
 #define E    (L/DIM)      // epsilon
 #define HBAR 1.054        // without the e-34
 
+
 #define MAX(a,b) a > b ? a : b
 #define MIN(a,b) a < b ? a : b
 #define ERROR(msg) do {                                               \
@@ -17,120 +18,13 @@
         } while(0);
 
 
-typedef struct _wave_function {
-    int n;             // the energy level of the wave function
-    float energy;      // the energy of the particle at that level
-    float probability; // the probability of getting this wave function (c^2_n)
-} _wave_function;
-
 typedef struct particle {
-    float mass;            // the mass of the particle
-    int waveCount;
+    float mass;
+    int energy_levels;
+    float *probabilities;
 } particle;
 
 
-
-/**
- * Create a new wave
- *
- * @param _wave_function  the wave function to create
- * @param int             the energy level of the particle
- * @param float           the mass of the particle
- * @param float           the probability of the wave function (c^2_n)
-*/
-void _create_wave(_wave_function *wave, int n, float mass, float p) {
-    wave->n = n;
-    wave->energy = HBAR*HBAR * PI*PI * n*n / (mass * L*L);
-    wave->probability = p;
-}
-
-
-/**
- * Given numbers, the function generates the next set of probabilities
- * based on these numbers.
- *
- * @param int            the numbers to generate the probabilities from
- * @param int            the length of the numbers array (should be the same
-                         for the probabilities array);
- * @param _wave_function array of wave function to update their probabilities
-*/
-void next_probabilities(int *numbers, int N, _wave_function** wave) {
-    int sum = 0;
-    int carry = 1;
-    for(int i=N-1;i>=0;i--) {
-        numbers[i] +=  carry;
-        carry       =  numbers[i] / 10;
-        numbers[i] %=  10;
-        sum        +=  numbers[i];
-    }
-    for(int i=0;i<N;i++) {
-        wave[i]->probability = 1.0f*numbers[i] / sum;
-    }
-}
-
-/**
- * Create a new particle
- * 
- * @param particle      the particle to create
- * @param int           the number of energy levels
- * @param float         the mass of the particle
-*/
[…]

### 2013-08-21 src/pbox.cu
Added documentation to the source code

@@ -3,7 +3,7 @@
 #include <math.h>
 
 
-#define PI   3.1415
+#define PI   3.141592653589793
 #define L    1.0          // the length of the box
 #define DIM  512          // the dimensions of the window (should be 2^n)
 #define E    (L/DIM)      // epsilon
@@ -18,13 +18,22 @@
         } while(0);
 
 
+/* The struct that describes a particle */
 typedef struct particle {
-    float mass;
-    int energy_levels;
-    float *probabilities;
+    float mass;            // mass of the particle
+    int energy_levels;     // the number of energy levels
+    float *probabilities;  // the probabilitis of every energy level
 } particle;
 
 
+/**
+ * CUDA kernel that finds the maximum in buckets of numbers
+ * and fills it in the partialMax array using the reduction method
+ *
+ * @param float*            the numbers to find the maximum of
+ * @param int               the length of the array (param 1)
+ * @param float*            list of buckets containing the maximums
+*/
 __global__
 void cuda_max(float *numbers, int N, float *partialMax) {
     extern __shared__ float cache[];
@@ -53,6 +62,14 @@ void cuda_max(float *numbers, int N, float *partialMax) {
         partialMax[blockIdx.x] = cache[0];
 }
 
+/**
+ * Function that finds the maximum of an array of numbers using the
+ * cuda_max function
+ *
+ * @param float*            the array of numbers to find the maximum of
+ * @param int               the length of the array
+ * @return float            the maximum of the array
+*/
 float max(float *numbers, int N) {
     int tpb = 256; // threads per block
     int bpg = MIN(32, (N+tpb-1)/tpb); // blocks per grid
@@ -77,6 +94,14 @@ float max(float *numbers, int N) {
     return m;
 }
 
+/**
+ * Given an array of numbers, the function finds a next set
+ * of probabilities
+ *
+ * @param int*                the numbers to act upon
+ * @param int                 the length of the array
+ * @param float*              where to store the next set of probabilities
+*/
 void next_probabilities(int *numbers, int N, float* probabilities) {
     int sum = 0;
     int carry = 1;
@@ -91,6 +116,13 @@ void next_probabilities(int *numbers, int N, float* probabilities) {
     }
 }
 
+/**
+ * A function that creates a new particle
+ *
+ * @param particle*           the particle to create
+ * @param int                 the number of energy levels the particle can have
+ * @param float               the mass of the particle
+*/
 void create_particle(particle *p, int N, float mass) {
[…]

### 2013-08-21 src/pbox.cu
Added energy related functions

@@ -282,6 +282,27 @@ float max_probability(particle *p) {
     return probability;
 }
 
+/**
+ * A function that finds the energy of the particle at a precise energy level
+ *
+ * @param particle*                 the particle
+ * @param int                       the energy level
+*/
+float energy(particle *p, int n) {
+    return (HBAR * HBAR * PI * PI * n * n) / (p->mass * L * L);
+}
+
+/**
+ * A function that finds the highest energy the particle can reach
+ * This function is useful for color mapping
+ *
+ * @param particle*                 the particle
+ * @return float                    the highest energy
+*/
+float highest_energy(particle *p) {
+    return energy(p, p->energy_levels);
+}
+
 int main(int argc, const char *argv[]) { 
     particle p;
     float elapsed;

### 2013-08-21 src/pbox.cu
Added an openGL sample and wrote some extra info in the paper

@@ -2,6 +2,10 @@
 #include <cuda.h>
 #include <math.h>
 
+#define GL_GLEXT_PROTOTYPES
+#include "GL/glut.h"
+#include "cuda_gl_interop.h"
+
 
 #define PI   3.141592653589793
 #define L    1.0          // the length of the box
@@ -16,6 +20,10 @@
             fprintf(stderr, "%s(%d): %s\n", __FILE__, __LINE__, msg); \
             exit(0);                                                  \
         } while(0);
+
+
+GLuint buffer;
+cudaGraphicsResource *resource;
 
 
 /* The struct that describes a particle */
@@ -180,8 +188,8 @@ void cuda_probability_to_map(float *probabilities, int n, float *map) {
     int offset = gridDim.x * blockDim.x;
 
     while(i < DIM * DIM) {
-        float x = 1.0*((int) (i/DIM))/DIM;
-        float y = 1.0*((int) (i%DIM))/DIM;
+        float x = 1.0*L*((int) (i%DIM))/DIM;
+        float y = 1.0*L*((int) (i/DIM))/DIM;
 
         if(x > 0 && x < L && y > 0 && y < L)
             map[i] = cuda_probability_2d_device(probabilities, n, x, y);
@@ -254,7 +262,6 @@ float probability(particle *p, float x, float y) {
  * @return                           the highest probability
 */
 float max_probability(particle *p) {
-    /*
      // Before analyzing the mathematical equation I would brute-force the problem
      // by looping over all the values and picking the maximum
     float *devProbabilities, *devMap;
@@ -269,9 +276,8 @@ float max_probability(particle *p) {
     cudaMemcpy(map, devMap, DIM * DIM * sizeof(float), cudaMemcpyDeviceToHost);
 
     return max(map, DIM*DIM);
-    */
 
-    float probability = 0.0, p_1d = 0.0;
+    float probability = E/L, p_1d = 0.0;
     int n;
     for(int i=0;i<10;i++) {
         n = i+1;
@@ -285,51 +291,113 @@ float max_probability(particle *p) {
 /**
  * A function that finds the energy of the particle at a precise energy level
  *
- * @param particle*                 the particle
+ * @param float                     the mass of the particle
[…]

### 2013-08-22 src/pbox.cu
Removed the OPENGL code to start a new one

@@ -113,7 +113,7 @@ float max(float *numbers, int N) {
 void next_probabilities(int *numbers, int N, float* probabilities) {
     int sum = 0;
     int carry = 1;
-    for(int i=N-1;i>=0;i--) {
+    for(int i=0;i<N;i++) {
         numbers[i] +=  carry;
         carry       =  numbers[i] / 10;
         numbers[i] %=  10;
@@ -311,93 +311,6 @@ float highest_energy(float mass, int n) {
     return energy(mass, n);
 }
 
-__global__
-void kernel(uchar4 *ptr, float *probabilities, int N, float max_proba) {
-    int i = threadIdx.x + blockIdx.x * blockDim.x;
-    int offset = blockDim.x * gridDim.x;
-
-    while(i < DIM*DIM) {
-        float x = L*((int) (i%DIM))/DIM;
-        float y = L*((int) (i/DIM))/DIM;
-        float p = cuda_probability_2d_device(probabilities, N, x, y)/max_proba;
-        float e = 0;
-        for(int j=0;j<N;j++) e += probabilities[j]*energy(1.0, j+1);
-        e /= highest_energy(1.0, N);
-
-        ptr[i].x = 255*p*e;
-        ptr[i].y = 20*p;
-        ptr[i].z = 255*p*(1-e);
-
-        i += offset;
-    }
-}
-
-void draw_func(void) {
-    glDrawPixels(DIM, DIM, GL_RGBA, GL_UNSIGNED_BYTE, 0);
-    glutSwapBuffers();
-}
-void key_func(unsigned char key, int x, int y) {
-    switch(key) {
-        case 27:
-            cudaGraphicsUnregisterResource(resource);
-            glBindBuffer(GL_PIXEL_UNPACK_BUFFER_ARB, 0);
-            glDeleteBuffers(1, &buffer);
-            exit(0);
-    }
-}
-
 int main(int argc, char *argv[]) { 
-    cudaDeviceProp prop;
-    int dev;
-
-    memset(&prop, 0, sizeof(cudaDeviceProp));
-    prop.major = 1;
-    prop.minor = 0;
-
-    cudaChooseDevice(&dev, &prop);
-    cudaGLSetGLDevice(dev);
-
-    glutInit(&argc, argv);
-    glutInitDisplayMode(GLUT_RGBA);
-    glutInitWindowSize(DIM, DIM);
[…]

### 2013-08-22 paper/paper.tex
Added the implementation text to the paper

--- 2013-08-16 paper/paper.tex
@@
 \section{Implementation}
-    \subsection{Memory Management}
-    \subsection{Optimizations}
-    \subsection{Implementations Details}
+    \subsection{Synopsis}
+    This simulation will display the probability of finding the particle in sub-boxes in the box
+    as well as display the energy in the sub-box of width $\epsilon$. Each frame displays the
+    particle with a new set of probabilities for each energy level.\\\\
+    The probability of the position will be visualized by the intensity of the color. The brighter
+    the color, the higher the probability.\\\\
+    The energy however will be visualized using the standard colors assigned to energies; lower
+    energies have blue-shades while high energies have red-shades.
+
+    \subsection{Conventions}
+    Some of the conventions adopted throughout the code are
+    \begin{itemize}
+        \item words in functions, structs and variable names are separated by an \_ (underscore)
+        \item CUDA kernels start with the keyword \texttt{cuda\_}.\\
+              For example \texttt{cuda\_probability}
+        \item CUDA device functions start with \texttt{cuda\_} and end with \texttt{\_device}.\\
+              For example \texttt{cuda\_probability\_1d\_device}
+        \item Tiny mathematical and physical constants, such as $\hbar$ are expressed
+              as float numbers without their orders.\\ For example $\hbar = 1.054\cdot10^{-34}$
+              is defined as \texttt{\#define HBAR 1.054}
+    \end{itemize}
+
+    \subsection{Brightness and Color mapping}
+        \subsubsection{Brightness}
+        The intensity of the colors denote the probability of finding the particle. Since
+        probabilities are always between 0 and 1, converting them to intensities is just
+        a matter of mapping the probabilities from $[0, 1]$ to $[0, 255]$.\\\\
+        However $\epsilon$ is a small number, so the probability is always small; usually
+        less than $0.1\%$. Therefore we need to find the highest probability in the space
+        and remap from $[0, \text{max}]$ to $[0, 255]$. At first I have tried to brute force
+        the problem exploiting my GPU using functions to find maximums using reduction. But
+        the process of finding the highest probability of a particle with 10 energy levels
+        inside a box divided into 262,144 boxes took around 28ms. Allowing me to get only
+        35fps in the animation.\\
+        A new solution needed to be adopted and I went over the equations again to try and
+        find that maximum analytically. My results are:\\\\
+        For $p(\alpha)$ to be maximal
+        $$\frac{\epsilon}{L} - \frac{1}{n_{\alpha}\pi} \sin(\frac{n_{\alpha}\pi}{L}\epsilon)
+        \cos{2\frac{n_{\alpha}\pi}{L}\alpha}$$
+        needs to be maximal.  And this is achieved when $\cos{x} = -1$ which is possible
+        since $0 \leq x \leq 2\pi$. Therefore the highest probability for one energy
+        level is: 
+        $$\frac{\epsilon}{L} + \frac{1}{n_{\alpha}\pi} \sin(\frac{n_{\alpha}\pi}{L}\epsilon)$$
+        Following similar analysis we can deduce that the highest probability (denoted by
+        $m$) is close to,
+        but not exactly:
+        $$m = m_x . m_y$$
+        $$m_{\alpha} = \frac{\epsilon}{L} + \sum_i \left ( c^2_i \frac{1}{n_{\alpha_i}\pi}
+        \sin\left(\frac{n_{\alpha_i}\pi}{L}\epsilon\right) \right )$$\\
+
+        \subsubsection{Tint and Color}
+        The tint in the color represents the energy of the particle. Red is the highest
+        energy and blue is the lowest. If we can map the energy in an interval between
+        0 and 1 we can easily get the RGB values. The following equation can be applied
+        $$ \begin{bmatrix} R_{xy} \\ G_{xy} \\ B_{xy} \end{bmatrix} = P(x, y)
+        \begin{bmatrix} 255E \\ 20 \\ 255(1-E) \end{bmatrix} $$
+        Where $P(x, y)$ is the probability of finding the particle at $(x, y)$ and between
+        0 and 1. And $E$ is the energy remapped between 0 and 1. \\\\
+        To problem of remapping the energy is easy to solve. Through mathematical analysis
+        we can show that the highest energy we can find is
+        $$ E_{max} = (n^2_{max_x} + n^2_{max_y})\frac{\hbar^2\pi^2}{2mL^2} $$
+        And therefore we can remap any energy we find to a value in the interval $[0, 1]$
+        and find the corresponding RGB values.
+
+\newpage
+\section{API}
 
@@
 \newpage
-\section{Code}
-    \subsection{Code}
-\begin{verbatim}
-\end{verbatim}
-    \subsection{License: GNU General Public License}
+\section{Code License: GNU General Public License}
 Copyright (C) 2013 George Zakhour\\

### 2013-08-23 src/pbox.cu
Fixed the code

@@ -106,6 +106,10 @@ float max(float *numbers, int N) {
  * Given an array of numbers, the function finds a next set
  * of probabilities
  *
+ * TODO Probabilities can be expressed as sine waves such that
+ *      the period of a wave is set such as every sub wave can
+ *      occur
+ *
  * @param int*                the numbers to act upon
  * @param int                 the length of the array
  * @param float*              where to store the next set of probabilities
@@ -113,7 +117,7 @@ float max(float *numbers, int N) {
 void next_probabilities(int *numbers, int N, float* probabilities) {
     int sum = 0;
     int carry = 1;
-    for(int i=0;i<N;i++) {
+    for(int i=N-1;i>=0;i--) {
         numbers[i] +=  carry;
         carry       =  numbers[i] / 10;
         numbers[i] %=  10;
@@ -262,8 +266,6 @@ float probability(particle *p, float x, float y) {
  * @return                           the highest probability
 */
 float max_probability(particle *p) {
-     // Before analyzing the mathematical equation I would brute-force the problem
-     // by looping over all the values and picking the maximum
     float *devProbabilities, *devMap;
     float *map;
 
@@ -277,6 +279,9 @@ float max_probability(particle *p) {
 
     return max(map, DIM*DIM);
 
+    /*
+    // This algorithm failed for many waves
+    // I had to fall back to the original bruteforce algorithm written above
     float probability = E/L, p_1d = 0.0;
     int n;
     for(int i=0;i<10;i++) {
@@ -286,6 +291,7 @@ float max_probability(particle *p) {
     }
 
     return probability;
+    */
 }
 
 /**
@@ -311,6 +317,96 @@ float highest_energy(float mass, int n) {
     return energy(mass, n);
 }
 
+__global__
+void kernel(uchar4 *ptr, float *probabilities, int N, float max_proba) {
+    int i = threadIdx.x + blockIdx.x * blockDim.x;
+    int offset = blockDim.x * gridDim.x;
+
+    while(i < DIM*DIM) {
+        float x = L*((int) (i%DIM))/DIM;
[…]

### 2013-08-23 paper/paper.tex
Fixed the code
Improved the next probability function and added the implementation description in the paper about the maximum proba and the next probability

@@
         For $p(\alpha)$ to be maximal
-        $$\frac{\epsilon}{L} - \frac{1}{n_{\alpha}\pi} \sin(\frac{n_{\alpha}\pi}{L}\epsilon)
-        \cos{2\frac{n_{\alpha}\pi}{L}\alpha}$$
+        $$\frac{\epsilon}{L} - \frac{1}{n_{\alpha}\pi} \sin\left(\frac{n_{\alpha}\pi}{L}\epsilon\right)
+        \cos\left(2\frac{n_{\alpha}\pi}{L}\alpha\right)$$
         needs to be maximal.  And this is achieved when $\cos{x} = -1$ which is possible
@@
         $$m_{\alpha} = \frac{\epsilon}{L} + \sum_i \left ( c^2_i \frac{1}{n_{\alpha_i}\pi}
-        \sin\left(\frac{n_{\alpha_i}\pi}{L}\epsilon\right) \right )$$\\
+        \sin\left(\frac{n_{\alpha_i}\pi}{L}\epsilon\right) \right )$$\\\\
+        \textbf{UPDATE} After implementing and testing the algorithm the uncertainty in the algorithm
+        has proved to be overwhelming and the probabilities' range was no longer $[0, 1]$ but much smaller.
+        Therefore a fallback was needed.\\\\
+        The approach followed was a bruteforce solution where we compute the probabilities at each at each
+        pixel and search for the largest. However to make the searching faster a reduction algorithm
+        was adopted to speed it up. The algorithm creates a much smaller array in which it stores
+        the maximum of a chunk of the original array. Then on the CPU we loop over the results (which are
+        usually 32) and pick the maximum of these. The following illustration describes how the maximum
+        reduction algorithm works
+        \begin{figure}[hb]
+            \centering
+                \includegraphics[width=8cm]{graphics/max_reduction.png}
+            \caption{Finding the maximum element in an array using reduction}
+        \end{figure}
+        The fact that this algorithm was used adds a restriction on the dimensions of the window. The height
+        and the width of the window need to be a power of 2 for this algorithm to work effectively.
+        No solution has been made up yet to generalize the dimensions of the window.
 
@@
         and find the corresponding RGB values.
+
+        \subsection{Next set of probabilities}
+        Every frame in the animation consists of a new set of probabilities close to the ones of the
+        previous frame. A problem came up and that is how to generate new set of probabilities close
+        to the ones in the previous frame to insure the smooth animation between the frames.\\\\
+        The model adopted to insure the smoothness and the continuity in the probabilities consists
+        of multiple sine waves where the period of one is 10 times more than its neighbor and so on.
+        If plotted, the model looks like the following: (the sines were stretched vertically for a better
+        visualization)
+        \begin{figure}[hb]
+            \centering
+                \includegraphics[width=8cm]{graphics/sine_probabilities.png}
+            \caption{Plotting
+               $x(t) = 10\left| sin \left( \frac{0.1}{\pi}t \right) \right|$, 
+               $y(t) = 10\left| sin \left( \frac{1}{\pi}t \right) \right|$, 
+               $z(t) = 10\left| sin \left( \frac{10}{\pi}t \right) \right|$}
+        \end{figure}
+        After finding the values of each sine wave at a time $t$ we can normalize the answers so that
+        their sum is $1$. Bellow is a table representing some values normalized.
+        \begin{center}
+            \begin{tabular}{| c | c | c | c |}
+                \hline
+                \textbf{Time} & \textbf{x} & \textbf{y} & \textbf{z} \\ \hline
+                0.1 & 0.0091 & 0.0915 & 0.8994 \\
+                0.2 & 0.0096 & 0.0957 & 0.8947 \\
+                0.3 & 0.0104 & 0.1035 & 0.8861 \\
+                0.4 & 0.0116 & 0.1159 & 0.8725 \\
+                0.5 & 0.0136 & 0.1350 & 0.8515 \\
+                0.6 & 0.0166 & 0.1648 & 0.8186 \\
+                0.7 & 0.0215 & 0.2135 & 0.7649 \\
+                0.8 & 0.0304 & 0.3006 & 0.6690 \\
+                0.9 & 0.0490 & 0.4834 & 0.4675 \\
+                1.0 & 0.0824 & 0.8102 & 0.1074 \\
+                1.1 & 0.0479 & 0.4698 & 0.4822 \\
+                1.2 & 0.0368 & 0.3590 & 0.6042 \\
+                1.3 & 0.0322 & 0.3134 & 0.6544 \\
+                1.4 & 0.0309 & 0.2987 & 0.6704 \\
+                1.5 & 0.0317 & 0.3053 & 0.6630 \\
+                1.6 & 0.0347 & 0.3324 & 0.6329 \\
+                1.7 & 0.0405 & 0.3859 & 0.5736 \\
+                1.8 & 0.0509 & 0.4818 & 0.4673 \\
+                1.9 & 0.0701 & 0.6595 & 0.2704 \\
+                2.0 & 0.0859 & 0.8022 & 0.1119 \\ \hline
+            \end{tabular}
+        \end{center}
+        \begin{center}Table 1: Values from $x(t)$, $y(t)$ and $z(t)$ at different times\end{center}
+        A general formula can be deduced to find the equation of the nth wave out of $N$ waves.
+        \begin{align}
+            P_{n/N}(t) = \left| sin\left( \frac{10^{2-(N-n)}}{\pi}t \right)\right|
+        \end{align}
 

### 2013-08-23 src/pbox.cu

@@ -103,28 +103,22 @@ float max(float *numbers, int N) {
 }
 
 /**
- * Given an array of numbers, the function finds a next set
- * of probabilities
+ * The set of probabilities can be expressed as a bunch of sine waves
+ * For more information read section 4.4 (Next set of probabilities) on
+ * page 9 in the paper
  *
- * TODO Probabilities can be expressed as sine waves such that
- *      the period of a wave is set such as every sub wave can
- *      occur
- *
- * @param int*                the numbers to act upon
+ * @param float               the time
  * @param int                 the length of the array
  * @param float*              where to store the next set of probabilities
 */
-void next_probabilities(int *numbers, int N, float* probabilities) {
-    int sum = 0;
-    int carry = 1;
-    for(int i=N-1;i>=0;i--) {
-        numbers[i] +=  carry;
-        carry       =  numbers[i] / 10;
-        numbers[i] %=  10;
-        sum        +=  numbers[i];
+void next_probabilities(float t, int N, float* probabilities) {
+    float sum = 0;
+    for(int i=0;i<N;i++) {
+        probabilities[i] = abs(sin(pow(10, 1-i)*t));
+        sum += probabilities[i];
     }
     for(int i=0;i<N;i++) {
-        probabilities[i] = 1.0f*numbers[i] / sum;
+        probabilities[i] /= sum;
     }
 }
 
@@ -383,13 +377,9 @@ int main(int argc, char *argv[]) {
 
     int N = atoi(argv[1]);
     particle *p;
-    int *numbers;
     p = (particle*) malloc(sizeof(particle));
     create_particle(p, 10, 1.5);
-    numbers = (int*) malloc(p->energy_levels*sizeof(int));
-    for(int i=0;i<p->energy_levels;i++) numbers[i] = 9*(i==0);
-    for(int i=0;i<N;i++)
-        next_probabilities(numbers, p->energy_levels, p->probabilities);
+    next_probabilities(10041.0, p->energy_levels, p->probabilities);
 
     float *devP;
 

### 2013-08-23 paper/paper.tex
Fixed a problem in the math of the probability sine waves

@@
         \begin{center}Table 1: Values from $x(t)$, $y(t)$ and $z(t)$ at different times\end{center}
-        A general formula can be deduced to find the equation of the nth wave out of $N$ waves.
+        A general formula can be deduced to find the equation of the nth wave.
         \begin{align}
-            P_{n/N}(t) = \left| sin\left( \frac{10^{2-(N-n)}}{\pi}t \right)\right|
+            P_{n}(t) = \left| sin\left( \frac{10^{1-n}}{\pi}t \right)\right|
         \end{align}

### 2013-08-23 src/pbox.cu

@@ -117,9 +117,8 @@ void next_probabilities(float t, int N, float* probabilities) {
         probabilities[i] = abs(sin(pow(10, 1-i)*t));
         sum += probabilities[i];
     }
-    for(int i=0;i<N;i++) {
-        probabilities[i] /= sum;
-    }
+    // Normalize the probabilities
+    for(int i=0;i<N;i++) probabilities[i] /= sum;
 }
 
 /**

### 2013-08-24 src/pbox.cu
Finished a working prototype

@@ -1,29 +1,25 @@
 #include <stdio.h>
 #include <cuda.h>
+#include <string.h>
 #include <math.h>
+#include <time.h>
 
 #define GL_GLEXT_PROTOTYPES
-#include "GL/glut.h"
-#include "cuda_gl_interop.h"
+#include <GL/glut.h>
+
+#include <cuda_runtime.h>
+#include <cuda_gl_interop.h>
 
 
 #define PI   3.141592653589793
-#define L    1.0          // the length of the box
-#define DIM  512          // the dimensions of the window (should be 2^n)
-#define E    (L/DIM)      // epsilon
-#define HBAR 1.054        // without the e-34
+#define L    1.0            // the length of the box
+#define DIM  512            // the dimensions of the window (should be 2^n)
+#define E    (L/DIM)        // epsilon
+#define HBAR 1.054          // without the e-34
 
 
 #define MAX(a,b) a > b ? a : b
 #define MIN(a,b) a < b ? a : b
-#define ERROR(msg) do {                                               \
-            fprintf(stderr, "%s(%d): %s\n", __FILE__, __LINE__, msg); \
-            exit(0);                                                  \
-        } while(0);
-
-
-GLuint buffer;
-cudaGraphicsResource *resource;
 
 
 /* The struct that describes a particle */
@@ -32,6 +28,37 @@ typedef struct particle {
     int energy_levels;     // the number of energy levels
     float *probabilities;  // the probabilitis of every energy level
 } particle;
+
+
+/** global variables used in the code */
+cudaGraphicsResource *resource;
+GLuint   buffer;
+float    INCREASE_TIME = 0.01;           // number to increase the time by
+particle p;                              // the particle
+float    t             = INCREASE_TIME;  // starting time
+int      frames        = 0;              // number of frames
+float    total_time    = 0.0f;           // total time
+int      PAUSE         = 0;              // whether the animation is paused
+
+
+/**
+ * Function headers
+*/
+float max(float*, int);
+void  next_probabilities(float, int, float*);
+void  create_particle(particle*, int, float);
[…]

### 2013-08-25 paper/paper.tex
Initial Commit (First version)

@@
 \hypersetup{colorlinks=false}
+\def\mar{\hspace*{5mm}}
 
@@
     \subsection{Brightness and Color mapping}
-        \subsubsection{Brightness}
+        \subsubsection{Brightness} \label{sec:brightness}
+        \label{sec:brightness}
         The intensity of the colors denote the probability of finding the particle. Since
@@
         \begin{figure}[hb]
+            \label{fig:reduction}
             \centering
@@
 
-        \subsubsection{Tint and Color}
+        \subsubsection{Tint and Color}\label{sec:colormap}
         The tint in the color represents the energy of the particle. Red is the highest
@@
         \subsection{Next set of probabilities}
+        \label{sec:nextproba}
         Every frame in the animation consists of a new set of probabilities close to the ones of the
@@
 
+    \subsection{(G)UI}
+    The GUI will be farely simple, a window where the simulation is drawn on. The user can
+    control the simulation by keystrokes listed bellow.
+    \begin{center}
+        \begin{tabular}{|r|r|}
+        \hline
+            . & Increase the time offset \\ \hline
+            , & Decrease the time offset \\ \hline
+            SPACE & Toggle pausing \\ \hline
+            0 & Reset the animate (t = 0) \\ \hline
+            ESC & Quit the simulation \\ \hline
+        \end{tabular}
+    \end{center}
+    As well, the user can control the number of wave functions that are being simulated using
+    command line arguments as follow
+    \begin{verbatim}pbox n\end{verbatim}
+    Where \verb|n| is the number of wave functions the user wishes to simulate.
+
 \newpage
 \section{API}
+    \subsection{Structures}
+        \subsubsection{particle}
+        {\bf Members}
+        \begin{itemize}
+            \item \begin{verbatim}float mass\end{verbatim} The mass of the particle.
+            \item \begin{verbatim}int energy_levels\end{verbatim} The number
+                  of energy levels the particle has, i.e. the number of wave
+                  functions that describe the particle.
+            \item \begin{verbatim}float *particles\end{verbatim} The
+                  probabilities of each wave function. The length of the array
+                  this variable points to should be \verb|energy_levels|.
+        \end{itemize}
+
+    \vspace{1cm}
+    \subsection{Global Variables}
+    The program uses these global variables
+    \begin{itemize}
+        \item \verb|cudaGraphicsResource *resource|\\\mar- The graphics resource that is
+              linked to the OpenGL environment.
+        \item \verb|GLuint buffer|\\\mar- The buffer used by OpenGL.
+        \item \verb|float INCREASE_TIME|\\\mar- The time offset used to increase the time
+              between the frames.
+        \item \verb|particle p|\\\mar- The particle that the program is about.
+        \item \verb|float t|\\\mar- The current time (set as the offset).
+        \item \verb|int frames|\\\mar- The number of frames.
+        \item \verb|float total_time|\\\mar- The total time.
+        \item \verb|int PAUSE|\\\mar- Whether the animation is paused.
+    \end{itemize}
+
+    \newpage
+    \subsection{Functions}
+        \vspace{1cm}
+        \subsubsection{float max(float *numbers, int N)}
+        This function finds the maximum number in an array of numbers using reduction
+        as described in ~\ref{fig:reduction}.\\
+        \\{\bf Parameters}\\
+        \verb|float *numbers|\\\mar- A pointer of the array to find the maximum of.\\
+        \verb|int N|\\\mar- The length of \verb|numbers|. {\bf Should be a power of 2}.\\
+        \\{\bf Returns}\\
+        \verb|float|, the maximum number in the array.
+
+        \vspace{1cm}
+        \subsubsection{void next\_probabilities(float t, int N, float *probabilities)}
+        This function generates a new set of probabilities as described in ~\ref{sec:nextproba}.
+        \\\\{\bf Parameters}\\
+        \verb|float t|\\\mar- The time parameter required in the algorithm.\\
+        \verb|int N|\\\mar- the number of parameters to generate, i.e. the length of
+        the next parameter.\\
+        \verb|float *probabilities|\\\mar- The array to write the new probabilities to.
+
+        \vspace{1cm}
+        \subsubsection{void create\_particle(particle *p, int N, float mass)}
+        This function creates a new particle.\\
+        \\{\bf Parameters}\\
+        \verb|particle *p|\\\mar- The particle to store the new particle in.\\
+        \verb|int N|\\\mar- The number of wave functions (energy levels) that describe
+        the particle.\\
+        \verb|float mass|\\\mar- The mass of the particle
+
+        \vspace{1cm}
+        \subsubsection{float probability(particle *p, float x, float y)}
+        This function finds the probability of finding a particle inside a box of center
+        $(x, y)$ and of width and height $\epsilon$.\\
+        \\{\bf Parameters}\\
+        \verb|particle *p|\\\mar- The particle to act on\\
+        \verb|float x|\\\mar- The x-coordinate of the particle\\
+        \verb|float y|\\\mar- The y-coordinate of the particle\\
+        \\{\bf Returns}\\
+        \verb|float|, the probability of finding the particle at $(x, y)$
+
+        \vspace{1cm}
+        \subsubsection{float max\_probability(particle *p)}
+        This function finds the highest probability of existing at a position $(x, y)$
+        in the space, i.e. inside the box. The function is used to later on map the
+        probabilities from $[0, \text{max}]$ to $[0, 255]$ for color-brightness purposes.
+        The algorithm adopted is described in ~\ref{sec:brightness}.\\
+        \\{\bf Parameters}\\
+        \verb|particle *p|\\\mar- The particle to find the highest probability of existing\\
+        \\{\bf Returns}\\
+        \verb|float|, the highest probability to exist.
+
+        \vspace{1cm}
+        \subsubsection{void initGL(int *argc, char **argv)}
+        This function initializes the OpenGL environment.\\
+        \\{\bf Parameters}\\
+        \verb|int *argc|\\\mar- The length of the next parameter.\\
+        \verb|char **argv|\\\mar- The parameters supplied to the OpenGL environment.
+
+        \vspace{1cm}
+        \subsubsection{void display()}
+        This function takes care of drawing the output on the canvas on every iteration.
+        As well it prints out the current time, the average time per frame in milliseconds
+        and the offset that is used to increase the time.
+
+        \vspace{1cm}
+        \subsubsection{void key(unsigned char k, int x, int y)}
+        This function manages the keystrokes in the canvas.\\
+        \\{\bf Parameters}\\
+        \verb|unsigned char k|\\\mar- The character pressed.\\
+        \verb|int x|\\\mar- The x-coordinate where the character is pressed.\\
+        \verb|int y|\\\mar- The y-coordinate where the character is pressed.
+
+        \vspace{1cm}
+        \subsubsection{void free\_resources()}
+        This function is used to clean after closing the window, i.e. free the resources.
+
+        \vspace{1cm}
+        \subsubsection{void createVBO(GLUint *buffer, cudaGraphicsResource **resource,
+            unsigned int flags)}
+        This function initializes the buffer and the resources that are used by OpenGL.\\
+        \\{\bf Parameters}\\
+        \verb|GLuint *buffer|\\\mar- The buffer used by OpenGL.\\
+        \verb|cudaGraphicsResource **resource|\\\mar- The CUDA resource to link to the buffer\\
+        \verb|unsigned int flags|\\\mar- The flags used by OpenGL.
+
+        \vspace{1cm}
+        \subsubsection{void launch\_kernel(uchar4 *pos)}
+        This function launches the kernel that fills the CUDA resource which will be used
+        to draw on the canvas.\\
+        \\{\bf Parameters}\\
+        \verb|uchar4 *pos|\\\mar- The array of pixel data
+
+        \vspace{1cm}
+        \subsubsection{void runCuda(cudaGraphicsResource **resource)}
+        This function creates the resources for the kernel and launches it.\\
+        \\{\bf Parameters}\\
+        \verb|cudaGraphicsResource **resource|\\\mar- The CUDA resource.
+
+        \vspace{1cm}
+        \subsubsection{void run(int argc, char **argv)}
+        This function runs everything, i.e. initializes the environment, selects
+        a valid GPU card, creates the resources and launches the GUI.\\
+        \\{\bf Parameters}\\
+        \verb|int argc| the length of the next parameter.\\
+        \verb|char **argc| the parameters used by the OpenGL environment.\\
+
+        \vspace{1cm}
+        \subsubsection{void usage(char* program\_name)}
+        This function prints out the help message.\\
+        \\{\bf Parameters}\\
+        \verb|char* program_name|\\\mar- The name of the program to run
+
+    \newpage
+    \subsection{CUDA Kernels}
+        \vspace{1cm}
+        \subsubsection{\_\_global\_\_ void cuda\_max(float *numbers, int N, float *partialMax)}
+        This kernel finds the maximum in buckets (sub-array) of the numbers array and
+        fills the partialMax array using the reduction method described in ~\ref{fig:reduction}.
+        \\\\{\bf Parameters}\\
+        \verb|float *numbers|\\\mar- The numbers to find the maximum of.\\
+        \verb|int N|\\\mar- The length of the array of numbers.\\
+        \verb|float *partialMax|\\\mar- The list containing the maximum of the buckets.
+
+        \vspace{1cm}
+        \subsubsection{\_\_global\_\_ void cuda\_probability\_to\_map(float *probabilities, int n,
+        float *map)}
+        This kernel maps the coordinate array to the probability array.\\
+        \\{\bf Parameters}\\
+        \verb|float *probabilities|\\\mar- The probability set of each wave function.\\
+        \verb|int n|\\\mar- The number of wave functions.\\
+        \verb|float *map|\\\mar- The array that will be filled with probabilities
+
+        \vspace{1cm}
+        \subsubsection{\_\_global\_\_ void cuda\_probability(float *p, in N, float x, float y,
+        float *probability)}
+        This kernel finds the probability of finding the particle in a certain position.\\
+        \\{\bf Parameters}\\
+        \verb|float *p|\\\mar- The probability set of each wave function.\\
+        \verb|int N|\\\mar- The number of energy levels.\\
+        \verb|float x|\\\mar- The x-coordinate of the particle.\\
+        \verb|float y|\\\mar- The y-coordinate of the particle.\\
+        \verb|float *probability|\\\mar- The variable to write the probability to.
+
+        \vspace{1cm}
+        \subsubsection{\_\_global\_\_ void kernel(uchar4 *ptr, float *probabilities, int N,
+        float max\_proba)}
+        This kernel fills the pixel array with the corresponding colors to display them.\\
+        \\{\bf Parameters}\\
+        \verb|uchar4 *ptr|\\\mar- The array of pixels\\
+        \verb|float *probabilities|\\\mar- The array of probabilities of each wave function.\\
+        \verb|int N|\\\mar- The number of wave functions.\\
+        \verb|gloat max_proba|\\\mar- The highest probability in the space.
+
+    \newpage
+    \subsection{CUDA Device Functions}
+        \vspace{1cm}
+        \subsubsection{\_\_device\_\_ float cuda\_probability\_1d\_device(int n, float x)}
+        This function finds the probability of finding the particle in one dimension 
+        at the position x and at the energy level n.\\
+        \\{\bf Parameters}\\
+        \verb|int n|\\\mar- The energy level of the particle\\
+        \verb|float x|\\\mar- The position of the particle\\
+        \\{\bf Returns}\\
+        \verb|float|, the probability.
+
+        \vspace{1cm}
+        \subsubsection{\_\_device\_\_ float cuda\_probability\_1d\_device(float *probabilities,
+        int n, float x, float y)}
+        This function finds the probability of finding the particle at a fixed position
+        given a set of probabilities, the number of energy levels and the position.\\
+        \\{\bf Parameters}\\
+        \verb|float *probability|\\\mar- The probability of each energy level.\\
+        \verb|int n|\\\mar- The number of energy levels.\\
+        \verb|float x|\\\mar- The x-coordinate of the particle.\\
+        \verb|float y|\\\mar- The y-coordinate of the particle.\\
+        \\{\bf Returns}\\
+        \verb|float|, the probability.
+
+        \vspace{1cm}
+        \subsubsection{\_\_device\_\_ float energy(float mass, int n)}
+        This function finds the energy of the particle at a precise energy level.\\
+        \\{\bf Parameters}\\
+        \verb|float mass|\\\mar- The mass of the particle.\\
+        \verb|int n|\\\mar- The energy level the particle is at.\\
+        \\{\bf Returns}\\
+        \verb|float|, the energy at the energy level n.
+
+        \vspace{1cm}
+        \subsubsection{\_\_device float highest\_energy(float mass, int n)}
+        This function finds the highest energy the particle can reach. This function
+        is used for color mapping described in ~\ref{sec:colormap}.\\
+        \\{\bf Parameters}\\
+        \verb|float mass|\\\mar- The mass of the particle.\\
+        \verb|int n|\\\mar- The highest energy level the particle can reach.\\
+        \\{\bf Returns}\\
+        \verb|float|, the highest energy
 
@@
 \section{Results}
+    \subsection{Screenshots}
+    These are some screenshots taken from the simulation. Only 5 wave functions are illustrated.\\
+    \begin{table}[ht]
+        \begin{center}
+        \begin{tabular}{c c c}
+            \includegraphics[width=3.5cm]{graphics/result1.png} &
+            \includegraphics[width=3.5cm]{graphics/result3.png} &
+            \includegraphics[width=3.5cm]{graphics/result2.png} \\
+            1 & 2 & 3 \\\\
+            \includegraphics[width=3.5cm]{graphics/result6.png} &
+            \includegraphics[width=3.5cm]{graphics/result4.png} &
+            \includegraphics[width=3.5cm]{graphics/result5.png} \\
+            4 & 5 & 6
+        \end{tabular}
+        \end{center}
+    \end{table}\\
+    From observing the screenshots we can generate a table that encapsulate the wave functions most visible
+    in the screenshots
+    \begin{table}[ht]
+        \begin{center}
+            \begin{tabular}{| r | c | c | c | c | c |}
+            \hline
+            Screenshot & 1 & 2 & 3 & 4 & 5 \\ \hline
+            1 & $\times$ & $\times$ & & & \\ \hline
+            2 & $\times$ & $\times$ & $\times$ & & $\times$ \\ \hline
+            3 & & $\times$ & $\times$ & $\times$ & \\ \hline
+            4 & & $\times$ & $\times$ & & $\times$ \\ \hline
+            5 & & & $\times$ & & $\times$ \\ \hline
+            6 & $\times$ & & & & $\times$ \\ \hline
+            \end{tabular}
+        \end{center}
+    \end{table}
+
+    \subsection{Benefits}
+    Thanks to this simulation the problem of the Particle in a box is now much clearer
+    to me, and I was able to understand more about the topic.\\\\
+    As well it helped me visualize another problem related and that is \emph{Heisenberg's
+    Uncertainty Principle} described mathematically as $\delta_x\delta_p \geq \frac{\hbar}{2}$.
+    As described in section ~\ref{sec:brightness} the more intense the color at a pixel is,
+    the higher the probability. For higher the quantum number is the larger the energy, and
+    since the energy is directly related to the momentum of the particle we should expect
+    a large uncertainty in the position of the particle. That is perfectly illustrated in this
+    simulation, as the graphics become red the higher the energy is and we can see numerous
+    bright spots on the canvas, meaning there is a high probability for the particle to exist
+    in many positions, therefore agreeing with Heisenberg's Uncertainty Principle.
 

### 2013-08-26 paper/paper.tex
Fixeed typos

@@
 \section{Background, Motive and Experiment}
-The particle in box problem is one of the first problems given to undergraduate
+The particle in a box problem is one of the first problems given to undergraduate
 students in a course about quantum mechanics. It showcases how the energy
@@
 The particle in a box is an experiment in which a particle is stuck inside a
-box and cannot escape. The energy outside this box is $\infty$ while inside it
-is $0$. The particle thus moves inside the box of dimensions $L$x$L$.\\\\
+box and cannot escape. The energy outside this box is $\infty$ while it
+is $0$ inside. The particle thus moves inside the box of dimensions $L$x$L$.\\\\
 Trying to imagine how these probabilities are scattered in the box is hard and one
@@
 (position, energy and momentum).\\\\
-In this simulation I try and simulate the "particle in a box" problem to
+In this simulation I try to simulate the "particle in a box" problem to
 display the probabilities of the position and the energy of the particle at
@@
         \end{equation}
-        The time independent Schrodinger equation is given as:
+        The time independent Schrodinger equation, which can be found from \ref{eq:time_dependant_equation}  is given as:
         \begin{equation} \label{eq:time_independent_equation}
@@
         Plugging (\ref{eq:suppose_twod_solution}) in (\ref{eq:time_independent_equation})
-        and doing similar operations as \emph{~\ref{sec:wave_equation_oned}} we get
+        and doing similar operations as ~\ref{sec:wave_equation_oned} we get
         the following solution:
@@
         \subsubsection{Energy in two dimensions}
-        Doing the same steps as \emph{~\ref{sec:energy_oned}} we find that the energy has
+        Doing the same steps as ~\ref{sec:energy_oned} we find that the energy has
         become:
@@
         \subsubsection{Probability function of the position}
-        Similar to section \emph{~\ref{sec:position_oned}} to find the
+        Similar to section ~\ref{sec:position_oned} to find the
         probability of finding the particle inside the box of size
@@
         $n_x$ and $n_y$ and describes the particle for these energy levels only.
-        However the final time-dependant equation is a combination of many of these
+        However the final time-independent equation is a combination of many of these
         equations. The final time-independent equation is:
@@
 \section{Implementation}
-    \subsection{Synopsis}
+    \subsection{Synopsis} \label{sec:synopsis}
     This simulation will display the probability of finding the particle in sub-boxes in the box
@@
         \item Tiny mathematical and physical constants, such as $\hbar$ are expressed
-              as float numbers without their orders.\\ For example $\hbar = 1.054\cdot10^{-34}$
+              as floating numbers without their orders.\\ For example $\hbar = 1.054\cdot10^{-34}$
               is defined as \texttt{\#define HBAR 1.054}
@@
         less than $0.1\%$. Therefore we need to find the highest probability in the space
-        and remap from $[0, \text{max}]$ to $[0, 255]$. At first I have tried to brute force
-        the problem exploiting my GPU using functions to find maximums using reduction. But
+        and remap from $[0, \text{max}]$ to $[0, 255]$. At first a bruteforce solution was adopted
+        exploiting my GPU to find maximums using reduction. But
         the process of finding the highest probability of a particle with 10 energy levels
-        inside a box divided into 262,144 boxes took around 28ms. Allowing me to get only
-        35fps in the animation.\\
-        A new solution needed to be adopted and I went over the equations again to try and
-        find that maximum analytically. My results are:\\\\
+        inside a box divided into 262,144 boxes took around 28ms.\\
+        A new solution needed to be adopted, going over the mathematics again, a solution came out:\\\\
         For $p(\alpha)$ to be maximal
@@
         Following similar analysis we can deduce that the highest probability (denoted by
-        $m$) is close to,
-        but not exactly:
+        $m$) is close to, but not exactly:
         $$m = m_x . m_y$$
@@
         Therefore a fallback was needed.\\\\
-        The approach followed was a bruteforce solution where we compute the probabilities at each at each
+        The approach followed was the bruteforce solution described above where we compute the probabilities at each
         pixel and search for the largest. However to make the searching faster a reduction algorithm
-        was adopted to speed it up. The algorithm creates a much smaller array in which it stores
+        was adopted. The algorithm creates a much smaller array in which it stores
         the maximum of a chunk of the original array. Then on the CPU we loop over the results (which are
         usually 32) and pick the maximum of these. The following illustration describes how the maximum
-        reduction algorithm works
+        reduction algorithm works for a sub-array.
         \begin{figure}[hb]
@@
         \end{figure}
-        The fact that this algorithm was used adds a restriction on the dimensions of the window. The height
+        The fact that this algorithm was used adds a restriction to the dimensions of the window. The height
         and the width of the window need to be a power of 2 for this algorithm to work effectively.
@@
         0 and 1. And $E$ is the energy remapped between 0 and 1. \\\\
-        To problem of remapping the energy is easy to solve. Through mathematical analysis
+        The problem of remapping the energy is easy to solve. Through mathematical analysis
         we can show that the highest energy we can find is
@@
         Every frame in the animation consists of a new set of probabilities close to the ones of the
-        previous frame. A problem came up and that is how to generate new set of probabilities close
-        to the ones in the previous frame to insure the smooth animation between the frames.\\\\
+        previous frame. A problem came up and it is how to generate a new set of probabilities close
+        to the ones in the previous frame given a time parameter to insure the smooth animation between the frames.\\\\
         The model adopted to insure the smoothness and the continuity in the probabilities consists
@@
 
-    \subsection{(G)UI}
-    The GUI will be farely simple, a window where the simulation is drawn on. The user can
-    control the simulation by keystrokes listed bellow.
+    \subsection{GUI and Configuration}
+    The GUI is simple; it consists of a window where the simulation is drawn on
+    as described in ~\ref{sec:synopsis}. The user can control the simulation using keystrokes listed bellow.
     \begin{center}
@@
         \hline
-            . & Increase the time offset \\ \hline
-            , & Decrease the time offset \\ \hline
+            . & Increase the time offset (Animation runs faster) \\ \hline
+            , & Decrease the time offset (Animation runs slower) \\ \hline
             SPACE & Toggle pausing \\ \hline
-            0 & Reset the animate (t = 0) \\ \hline
+            0 & Reset the animation (t = 0) \\ \hline
             ESC & Quit the simulation \\ \hline
@@
     \end{table}\\
-    From observing the screenshots we can generate a table that encapsulate the wave functions most visible
-    in the screenshots
+    From observing the screenshots we can generate a table that encapsulates the wave functions most visible.
     \begin{table}[ht]
@@
     to me, and I was able to understand more about the topic.\\\\
-    As well it helped me visualize another problem related and that is \emph{Heisenberg's
+    It helped me visualize another related problem as well, which is \emph{Heisenberg's
     Uncertainty Principle} described mathematically as $\delta_x\delta_p \geq \frac{\hbar}{2}$.
     As described in section ~\ref{sec:brightness} the more intense the color at a pixel is,
-    the higher the probability. For higher the quantum number is the larger the energy, and
+    the higher the probability. The higher the quantum number is, the larger the energy, and
     since the energy is directly related to the momentum of the particle we should expect

### 2013-08-26 src/pbox.cu

@@ -413,8 +413,9 @@ void display() {
     total_time += diff;
     frames++;
 
-    printf("%-5.3f (Average time per frame %.5f ms) (+%.3f)\r",
+    printf("%-5.3f (Average time per frame %.5f ms) (+%.3f)",
            t, total_time/frames, INCREASE_TIME);
+    for(int i=0;i<200;i++) printf("\r");
 
     cudaEventDestroy(start);
     cudaEventDestroy(stop);
@@ -528,14 +529,18 @@ void run(int argc, char **argv) {
 }
 
 
-
+/**
+ * Prints out the help message of the program
+ *
+ * @param char      the name of the program
+*/
 void usage(char* program_name) {
     printf("A particle in a box simulation\n");
     printf("Usage: %s n\n", program_name);
     printf("n\tThe number of energy levels to simulate (default is 5)\n\n");
     printf("Pressing these following keys will:\n");
-    printf(".\tIncrease the time delay\n");
-    printf(",\tDescrease the time delay\n");
+    printf(".\tIncrease the time offset\n");
+    printf(",\tDescrease the time offset\n");
     printf("<space>\tToggle pausing\n");
     printf("0\tReset the animation\n");
     printf("<esc>\tQuit\n\n");

### 2013-08-29 src/pbox.cu
Added a command line interface

@@ -3,6 +3,7 @@
 #include <string.h>
 #include <math.h>
 #include <time.h>
+#include <curses.h>
 
 #define GL_GLEXT_PROTOTYPES
 #include <GL/glut.h>
@@ -40,6 +41,10 @@ int      frames        = 0;              // number of frames
 float    total_time    = 0.0f;           // total time
 int      PAUSE         = 0;              // whether the animation is paused
 
+/** ncurses variables */
+WINDOW* window;
+int ncols, nrows;
+
 
 /**
  * Function headers
@@ -59,6 +64,9 @@ void  launch_kernel(uchar4);
 void  runCuda(cudaGraphicsResource **resource);
 void  run(int, char**);
 void  usage(char*);
+/** For ncurses */
+void  clear_row(int);
+void  cmd_display();
 
 
 /**
@@ -385,64 +393,34 @@ void initGL(int *argc, char **argv) {
 }
 
 /**
- * The display function that runs on every iteration
-*/
-void display() {
-    cudaEvent_t start, stop;
-    cudaEventCreate(&stop);
-    cudaEventCreate(&start);
-    cudaEventRecord(start, 0);
-    
-    glBindBuffer(GL_PIXEL_UNPACK_BUFFER_ARB, buffer);
-    glBufferData(GL_PIXEL_UNPACK_BUFFER_ARB, DIM*DIM*4, NULL, GL_DYNAMIC_DRAW_ARB);
-    cudaGraphicsGLRegisterBuffer(&resource, buffer, cudaGraphicsMapFlagsNone);
-    glClear(GL_COLOR_BUFFER_BIT);
-    runCuda(&resource);
-    glDrawPixels(DIM, DIM, GL_RGBA, GL_UNSIGNED_BYTE, 0);
-
-    glutSwapBuffers();
-    t += (1-PAUSE)*INCREASE_TIME;
-
-    cudaEventRecord(stop, 0);
-    cudaEventSynchronize(stop);
-    float diff;
-    cudaEventElapsedTime(&diff, start, stop);
-
-    glutPostRedisplay();
-
-    total_time += diff;
-    frames++;
-
-    printf("%-5.3f (Average time per frame %.5f ms) (+%.3f)",
-           t, total_time/frames, INCREASE_TIME);
-    for(int i=0;i<200;i++) printf("\r");
-
-    cudaEventDestroy(start);
-    cudaEventDestroy(stop);
-
-}
-
-/**
  * Manage key strokes
  *
  * @param unsigned char         the character pressed
[…]

### 2013-08-30 paper/paper.tex
Added the documentation and code for a Command line UI

@@
 
-    \subsection{GUI and Configuration}
+    \subsection{GUI, Command line UI and User Configuration}
     The GUI is simple; it consists of a window where the simulation is drawn on
@@
     \begin{center}
-        \begin{tabular}{|r|r|}
+        \begin{tabular}{|r|l|}
         \hline
@@
             , & Decrease the time offset (Animation runs slower) \\ \hline
+            n & Go back one frame \\ \hline
+            m & Go forward one frame\\ \hline
             SPACE & Toggle pausing \\ \hline
@@
     \begin{verbatim}pbox n\end{verbatim}
-    Where \verb|n| is the number of wave functions the user wishes to simulate.
+    Where \verb|n| is the number of wave functions the user wishes to simulate.\\\\
+    {\bf UDPATE} A command line UI is used to display in an ordered fashion the state of the simulation
+    (time, energy levels, average time per frame and the increment used by the animation), as well it
+    displays the probabilities for each wave function and keyboard shortcuts and how to use the application.\\
+    The command line UI was built using the \verb|curses| library available on Linux, and in the latest
+    release of Xcode but unavailable directly on Windows\footnote{PDCurses is an alternative suggested
+    by many http://pdcurses.sourceforge.net/. The author did not test it!}.\\
+    This is a screenshot of the command line UI.\\\\
+    \begin{figure}[hb]
+        \centering
+        \includegraphics[width=14cm]{graphics/cmdui.png}
+        \caption{Command-Line UI}
+    \end{figure}
 
@@
         \item \verb|int PAUSE|\\\mar- Whether the animation is paused.
+        \item \verb|WINDOW* window|\\\mar- The window variable used by \verb|ncurses| library.
+        \item \verb|int ncols, nrows|\\\mar- The number of cols and rows in the \verb|window| variable.
     \end{itemize}
@@
         \verb|char* program_name|\\\mar- The name of the program to run
+
+        \vspace{1cm}
+        \subsubsection{void clear\_row(int y)}
+        This function uses the \verb|ncuses| library to clear a row, i.e. to fill it with
+        empty characters with the background attitude.\\
+        \\{\bf Parameters}\\
+        \verb|int y|\\\mar- The row number to clear.
+
+        \vspace{1cm}
+        \subsubsection{void cmd\_display()}
+        This function creates and fills the Command line UI using the \verb|ncurses| library.
+
+        \vspace{1cm}
+        \subsubsection{void init\_curses()}
+        This function initiates the \verb|curses| environment and fills the variables \verb|window|,
+        \verb|nrows|, \verb|ncols|, creates the colors for the background and text messages.
 

### 2013-08-30 src/pbox.cu

@@ -7,7 +7,6 @@
 
 #define GL_GLEXT_PROTOTYPES
 #include <GL/glut.h>
-
 #include <cuda_runtime.h>
 #include <cuda_gl_interop.h>
 
@@ -67,6 +66,7 @@ void  usage(char*);
 /** For ncurses */
 void  clear_row(int);
 void  cmd_display();
+void  init_curses();
 
 
 /**
@@ -403,7 +403,6 @@ void key(unsigned char k, int x, int y) {
     switch(k) {
         case 27:
             free_resources();
-            printf("\n");
             endwin();
             exit(0);
         case '.':
@@ -418,9 +417,9 @@ void key(unsigned char k, int x, int y) {
                 glutPostRedisplay();
             } break;
         case 'm':
-            t += INCREASE_TIME; frames++; glutPostRedisplay(); break;
+            t += INCREASE_TIME; glutPostRedisplay(); break;
         case 'n':
-            t -= INCREASE_TIME; frames--; glutPostRedisplay(); break;
+            t -= INCREASE_TIME; glutPostRedisplay(); break;
     }
 }
 
@@ -535,10 +534,13 @@ void usage(char* program_name) {
     attroff(COLOR_PAIR(2));
     move(y+10, 0);
     attron(COLOR_PAIR(5));
-    printw("%s [n]\nDESCRIPTION\n  n\n\    The number of wave functions to simulate", program_name);
+    printw("%s [n]\nDESCRIPTION\n  n\n    The number of wave functions to simulate", program_name);
     attroff(COLOR_PAIR(5));
 }
 
+/**
+ * Function that clears the rows
+ */
 void clear_row(int y) {
     move(y, 0);
     for(int i=0;i<ncols;i++) {
@@ -549,6 +551,9 @@ void clear_row(int y) {
     move(y, 0);
 }
 
+/**
+ * Create the command line interface
+*/
 void cmd_display() {
     // Draw the title
     attron(COLOR_PAIR(2));
@@ -596,7 +601,6 @@ void cmd_display() {
     attroff(COLOR_PAIR(5));
 
     move(0, ncols);
-
     refresh();
 }
 
@@ -637,6 +641,10 @@ void display() {
     cmd_display();
 }
 
+/**
+ * Function that initializes the curses library
+ * Sets up colors and fills the background
+*/
 void init_curses() {
     /** Initialize the ncurses variables */
     window = initscr();
@@ -671,11 +679,12 @@ int main(int argc, char *argv[]) {
     int N = 5;
 
[…]
